---
title: 'Tidy Tuesday: Black lives Matter'
author: Montana Kay Lara
date: '2020-06-24'
slug: tidy-tuesday-black-lives-matter
categories:
  - TidyTuesday
tags:
  - tidyverse
---

This workflow uses the African American Achievements Tidy Tuesday dataset, downloaded June 24, 2020, for the purposes of learning and education on the intersection of racial matters and data science. 

## Outline
I. Setup
II. Data
III. Clean Data
IV. Look at Data
V. Plots

## I. Setup

```{r packages}
library(tidyverse)
library(ggplot2)
library(tidytuesdayR)

#you would need the following packages if you were going to run the cleaning script instead of the II.Data grabs
library(readr)
library(xml2)
library(rvest)
library(janitor)
library(purrr)
library(stringr)
library(tibble)
library(here)
library(magrittr)
```

## II. Data

Dataset provided here: https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-06-09 
```{r data_grab}
# Get the Data viar readr read_csv
firsts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-09/firsts.csv')
science <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-09/science.csv')

# Or read in with tidytuesdayR package (https://github.com/thebioengineer/tidytuesdayR)
# Install via devtools::install_github("thebioengineer/tidytuesdayR")
# Either ISO-8601 date or year/week works!
tuesdata <- tidytuesdayR::tt_load('2020-06-09')
tuesdata <- tidytuesdayR::tt_load(2020, week = 24)

#Get the two datasets out of the tuesdata
firsts <- tuesdata$firsts
science <- tuesdata$science
```


## III. Clean Data

Tidy tuesday provides their cleaning script used to make the two datasets - and it is worth at least looking at to see how they build these. 
```{r clean_it_up_SCI, eval=FALSE}
# URL 1
url_science <- "https://en.wikipedia.org/wiki/List_of_African-American_inventors_and_scientists"
raw_html_sci <- read_html(url_science)

get_source <- function(x){
  raw_html_sci %>% 
    html_nodes("tbody") %>% 
    .[[2]] %>% 
    html_nodes(glue::glue("tr:nth-child({x})")) %>% 
    html_nodes("td:nth-child(1) > a") %>% 
    html_attr("href")
}

raw_sci_tab <- raw_html_sci %>% 
  html_table(fill = TRUE) %>% 
  .[[2]] %>% 
  janitor::clean_names() %>% 
  as_tibble() %>% 
  mutate(links = map(row_number(), get_source))

clean_sci <- raw_sci_tab %>% 
  mutate(references = str_replace_all(references, "\\]", ","),
         references = str_remove_all(references, "\\[")) %>% 
  unnest(links) %>% 
  mutate(links = paste0("https://en.wikipedia.org", links)) %>% 
  separate(years, into = c("birth", "death"), sep = "–") %>% 
  mutate(across(c(birth, death), as.integer)) %>% 
  mutate(occupation_s = str_replace_all(occupation_s, ",", ";"))

clean_sci %>% 
  filter(str_detect(tolower(occupation_s), "statistician"))


sci_citations <- raw_html_sci %>% 
  html_node("#mw-content-text > div > div.reflist > div") %>% 
  html_nodes("li") %>% 
  html_text() %>% 
  str_remove("\\^ ") %>% 
  enframe() %>% 
  rename(citation_num = name, citation = value) %>% 
  mutate(citation = str_replace_all(citation, "\"", "'"),
         citation = str_remove_all(citation, "\\n"))

sci_citations

clean_sci %>%
  add_row(tibble(
    name = "Amos, Harold", birth = 1918, death = 2003, occupation_s = "Microbiologist",
    inventions_accomplishments = "First African-American department chair at Harvard Medical School",
    references = "6,", links = "https://en.wikipedia.org/wiki/Harold_Amos"), .before = 1
  )
  
write_csv(clean_sci, path = here("2020/2020-06-09/science.csv"))
science <- read_csv("2020/2020-06-09/science.csv")
```

```{r clean_it_up_FIRSTS, eval=FALSE}
first_url <- "https://en.wikipedia.org/wiki/List_of_African-American_firsts"
raw_first <- read_html(first_url)

get_year <- function(id_num) {
  raw_first %>% 
    html_nodes(glue::glue("#mw-content-text > div > h4:nth-child({id_num}) > span")) %>% 
    html_attr("id") %>% 
    .[!is.na(.)]
}

get_first <- function(id_num){
  raw_first %>% 
    html_nodes(glue::glue("#mw-content-text > div > ul:nth-child({id_num})")) %>% 
    html_text() %>% 
    str_split("\n")
}

tidyr::crossing(id_num = 9:389, count = 1:5)

raw_first_df <- tibble(id_num = 9:390) %>% 
  mutate(year = map(id_num, get_year),
         text = map(id_num, get_first))

clean_first <- raw_first_df %>% 
  mutate(year = as.integer(year)) %>% 
  fill(year) %>% 
  unnest(text) %>% 
  unnest(text) %>% 
  separate(text, into = c("role", "person"), sep = ": ") %>% 
  mutate(person = str_remove_all(person, "\\\\"),
         person = str_trim(person),
         role = str_replace(role, "African American", "African-American")) %>% 
  select(year, role, person)

clean_first %>% 
  group_by(year) %>% 
  summarize(n =n())

first_role <- function(category){
  str_detect(tolower(role), category)
}

edu <- paste0(c(
  "practice", "graduate", "learning", "college", "university", "medicine",
  "earn", "ph.d.", "professor", "teacher", "school", "nobel", "invent", "patent",
  "medicine", "degree", "doctor", "medical", "nurse", "physician", "m.d.", "b.a.", "b.s.", "m.b.a",
  "principal", "space", "astronaut"
), collapse = "|")

religion <- c("bishop", "rabbi", "minister", "church", "priest", "pastor", "missionary",
              "denomination", "jesus", "jesuits", "diocese", "buddhis") %>%
  paste0(collapse = "|")

politics <- c(
  "diplomat", "elected", "nominee", "supreme court", "legislature", "mayor", "governor",
  "vice President", "president", "representatives", "political", "department", "peace prize",
  "ambassador", "government", "white house", "postal", "federal", "union", "trade",
  "delegate", "alder", "solicitor", "senator", "intelligience", "combat", "commissioner",
  "state", "first lady", "cabinet", "advisor", "guard", "coast", "secretary", "senate",
  "house", "agency", "staff", "national committee"
) %>%
  paste0(collapse = "|")

sports <- c(
  "baseball", "football", "basketball", "hockey", "golf", "tennis",
  "championship", "boxing", "games", "medal", "game", "sport", "olympic", "nascar",
  "coach", "trophy", "nba", "nhl", "nfl", "mlb", "stanley cup", "jockey", "pga",
  "race", "driver", "ufc", "champion"
) %>%
  paste0(collapse = "|")

military <- c(
  "serve", "military", "enlist", "officer", "army", "marine", "naval",
  "officer", "captain", "command", "admiral", "prison", "navy", "general",
  "force"
) %>%
  paste0(collapse = "|")

law <- c("american bar", "lawyer", "police", "judge", "attorney", "law", 
         "agent", "fbi") %>%
  paste0(collapse = "|")

arts <- c(
  "opera", "sing", "perform", "music", "billboard", "oscar", "television",
  "movie", "network", "tony award", "paint", "author", "book", "academy award", "curator",
  "director", "publish", "novel", "grammy", "emmy", "smithsonian",
  "conduct", "picture", "pulitzer", "channel", "villain", "cartoon", "tv", "golden globe",
  "comic", "magazine", "superhero", "pulitzer", "dancer", "opry", "rock and roll", "radio",
  "record") %>%
  paste0(collapse = "|")

social <- c("community", "freemasons", "vote", "voting", "rights", "signature", 
            "royal", "ceo", "community", "movement", "invited", "greek", "million",
            "billion", "attendant", "chess", "pilot", "playboy", "own", "daughter",
            "coin", "dollar", "stamp", "niagara",
            "stock", "north pole", "reporter", "sail around the world", "press", "miss ",
            "everest")  %>%
  paste0(collapse = "|")

first_df <- clean_first %>% 
  mutate(gender = if_else(str_detect(role, "woman|Woman|her|she|female"), 
                          "Female African American Firsts", "African-American Firsts"),
         role = str_remove_all(role, "\""),
         person = str_remove_all(person, "\""),
         category = case_when(
           str_detect(tolower(role), military) ~ "Military",
           str_detect(tolower(role), law) ~ "Law",
           str_detect(tolower(role), arts) ~ "Arts & Entertainment",
           str_detect(tolower(role), social) ~ "Social & Jobs",
           str_detect(tolower(role), religion) ~ "Religion",
           str_detect(tolower(role), edu) ~ "Education & Science",
           str_detect(tolower(role), politics) ~ "Politics",
           str_detect(tolower(role), sports) ~ "Sports",
           TRUE ~ NA_character_
         )) %>% 
  rename(accomplishment = role)

first_df %>% write_csv(path = "2020/2020-06-09/firsts.csv")

firsts <- read_csv("2020/2020-06-09/firsts.csv")

plot_ex <- first_df %>% 
  mutate(n = 1) %>% 
  group_by(category) %>% 
  mutate(roll_n = cumsum(n)) %>% 
  ggplot(aes(x = year, y = roll_n, color = category)) +
  geom_step(size = 1) +
  theme(legend.position = "top") +
  tomtom::theme_tomtom() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  scale_x_continuous(breaks = seq(1750, 2020, 25)) +
  geom_hline(yintercept = 0, size = 1, color = "black") +
  labs(x = "", y = "",
       title = "Cumulative African-Americans firsts over time",
       subtitle = "479 'Firsts' of African-Americans breaking the color barrier across a wide range of topics",
       caption = "Data: wikipedia.org/wiki/List_of_African-American_firsts")

ggsave("2020/2020-06-09/pic2.png", plot_ex, height = 8, width = 14, units = "in", dpi = "retina")
```


## IV. Look at Data

```{r take_a_look}
#use head(), str(), class(), View(), etc. to see what the tables look like
head(science)
head(firsts)

```

## V. Plots

```{r plot}
science %<>% 
  filter(!is.na(death))
  
science %>% 
  ggplot(aes(x = birth, y = 0, label = name)) +
  geom_point()
```

